---
title: "Zimbi Dental All Vists"
author: "AlexMartinez"
date: "10/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Much like a regressions or classification model, there are certain parameters and hyperparameters that we can adjust to fit the best fitting model. Time series models have similar hyperparameters but have limited parameters that can be changed. The challenge in createing the best fit time series model comes from being able to check the error rate and check that the assumptions are correct.

### Data Cleaning

```{r}
library(tidyverse)
library(lubridate)
library(forecast)
library(zoo)

all_vists <- read.csv('Data/DentalIntelFiles/DentalIntelFiles_all_vists.csv')
all_vists_monthly <- all_vists %>% 
  mutate(
    startOfMonth = floor_date(as_date(mdy_hm(AppointmentDate, tz = NULL)), 'month')
  ) %>% 
  group_by(
    startOfMonth
  ) %>% 
  summarise(Appointments = n())



all_vists_monthly

```
We have cleaned that data so that each time period is one month and that each month has appointments. Next is to convert the tibble into a data type that is interpretable for time series analysis.

### Tibble to Time Series

```{r}
all_vists_monthly_ts <- all_vists_monthly %>%
  pull(Appointments) %>% 
  ts(start = c(2017,6), frequency = 12)

all_vists_monthly_ts
```
Interpretable Time Series Data Frame

### Time Series Visualization and Naive Forecast
```{r}
ggplot(data = all_vists_monthly) + 
  geom_line(aes(x = startOfMonth, y = Appointments))+
  geom_point(aes(x = startOfMonth, y = Appointments))
```

The naive forecast that I will use is a $t = 1$ forecast which will shift all values one position into the future.

```{r}
## T = 1

all_vists_monthly$t_1 <- as.vector(all_vists_monthly_ts) %>% lag()


```

```{r}

ggplot(data = all_vists_monthly) + 
  geom_line(aes(x = startOfMonth, y = Appointments), color = 'firebrick')+
  geom_point(aes(x = startOfMonth, y = Appointments), color = 'firebrick')+
  geom_line(aes(x = startOfMonth, y = t_1), color = 'skyblue3', alpha = 0.5)+
  geom_point(aes(x = startOfMonth, y = t_1), alpha = 0.5,color = 'skyblue3')

```

### Another Method

```{r}
nValid <- 48
nTrain <- length(all_vists_monthly_ts) - nValid
train.ts <- window(all_vists_monthly_ts, start = c(2017,6), end = c(2017,nValid))
valid.ts <- window(all_vists_monthly_ts, start = c(2017, nTrain + 1), end = c(2017, nTrain+nValid))
appointments.lm <- tslm(train.ts ~ trend + I(trend^2))
appointments.lm.pred <- forecast(appointments.lm, h = nValid, level = 0)

plot(appointments.lm.pred, ylim = c(0,200), ylab = 'Appointments', xlab = 'Time', bty = 'l',
xaxt = 'n', xlim = c(2017.5,2021), main = 'Zimbi Dental Appointment Forecast', flty = 2)
# plot(appointments.lm.pred, xlim = )
axis(1, at = seq(2017,2022,1), labels = format(seq(2017,2022,1)))
lines(appointments.lm$fitted, lwd = 3)
lines(valid.ts)

```

#### Evaluation

```{r}
accuracy(appointments.lm.pred$mean, valid.ts)
```

Provides various error metrics to compare. 

Next, we will find the various residuals of the model.

```{r}
names(appointments.lm.pred)
appointments.lm.pred$residuals
valid.ts - appointments.lm.pred$mean

hist(appointments.lm.pred$residuals, ylab = 'Frequency', xlab = 'Forecast Error')
```


#### Evaluation with Season and Non Season Naive Forecasting

```{r}
nValid <- 48
nTrain <- length(all_vists_monthly_ts) - nValid
train.ts <- window(all_vists_monthly_ts, start = c(2017,6), end = c(2017,nValid))
valid.ts <- window(all_vists_monthly_ts, start = c(2017, nTrain + 1), end = c(2017, nTrain+nValid))
naive.pred <- naive(train.ts, h = nValid)
snaive.pred <- snaive(train.ts, h = nValid)
accuracy(naive.pred, valid.ts)
accuracy(snaive.pred, valid.ts)

```


##### Moving Average Smoothing Method

```{r}
ma.trailing <- rollmean(all_vists_monthly_ts, k = 12, align = 'right')
ma.centered <- ma(all_vists_monthly_ts, order = 12)
plot(all_vists_monthly_ts, ylim = c(0,180), ylab = 'Appointments', xlab = 'Time', bty = 'l',
xaxt = 'n', xlim = c(2017.5,2021), main = 'Zimbi Dental Appointment Forecast')
axis(1, at = seq(2017,2022,1), labels = format(seq(2017,2022,1)))
lines(ma.centered, lwd = 2)
lines(ma.trailing, lwd = 2, lty = 2)
legend(2018, 180, c('Appointments','Centered Moving Average','Trailing Moving Average'), lty = c(1,1,2),
       lwd=c(1,2,2), bty = 'n')


```

##### Differencing Methods

Removing Trend (De-Trending) Lag 1 Difference
This type of differencing method removes the trend. 

```{r}

nValid <- 48
nTrain <- length(all_vists_monthly_ts) - nValid
train.ts <- window(all_vists_monthly_ts, start = c(2017,6), end = c(2017,nValid))
valid.ts <- window(all_vists_monthly_ts, start = c(2017, nTrain + 1), end = c(2017, nTrain+nValid))

diff(all_vists_monthly_ts, lag = 1)
plot(diff(all_vists_monthly_ts, lag = 1))
```


Removing Seasonality 
To remove seasonality, we find the difference according to the M interval that we are trying to adjust (in this interval M= 12). 

```{r}
diff(all_vists_monthly_ts, lag = 12)
plot(diff(all_vists_monthly_ts, lag = 12))
```

Removing Trend and Seasonality
Double remove each trend both trend (lag=1) and seasons seasonlity (lag=12)

```{r}
diff(diff(all_vists_monthly_ts, lag = 12), lag = 1)
plot(diff(diff(all_vists_monthly_ts, lag = 12), lag = 1))
```

Modeling with  Seasonality
```{r}
train.ts.season <- tslm(train.ts ~ season)
summary(train.ts.season)
```


Modeling with Seasonality and Trend
```{r}

train.ts.trend.season <- tslm(train.ts ~ trend + I(trend^2) + season)
summary(train.ts.trend.season)

```


Playing around with a model
```{r}
train.ts.trend.diff <- tslm(log(train.ts) ~ I(trend**2) + I((cos(trend)/12))+I(sin(2*pi*trend/12)))
summary(train.ts.trend.diff)
plot(train.ts.trend.diff$residuals)
```


Replacing Outliers

```{r}
fit <- nnetar(tsclean(train.ts))
fit.ts <- fit$fitted

train.ts.trend.diff <- tslm(log(fit.ts) ~ I(trend**2) + I((cos(trend)/12))+I(sin(2*pi*trend/12)))
summary(train.ts.trend.diff)
plot(train.ts.trend.diff$residuals)
```

Outlier Replaced

```{r}
train.ts.trend.season <- tslm(log(fit.ts) ~ trend + I(trend^2))
summary(train.ts.trend.season)

train.ts.trend.season <- tslm(train.ts ~ trend + I(trend^2))
summary(train.ts.trend.season)
```
