---
title: "TestWithTotalVisits"
author: "AlexMartinez"
date: "10/25/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

If you're reading this, then you are probably more excited about this that I am, or perhaps, you are looking to critique my work. Regardless, time series analysis is a fundamental tool in mathematics to help us understand trends, seasonality, and randomness in when looking at counts of an event occuring or looking at aggregations of numbers in a given time period. For this analysis we will look at dental office visits.

```{r LoadingLibrariesAndData, include=FALSE}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)

# Note: The Tutorial Uses Different Test Data

all_vists <- read.csv('../Data/DentalIntelFiles/DentalIntelFiles_all_vists.csv')
```

### Aggregating Data {.tabset}

```{r AggregatingData, collapse=TRUE}
all_vists_monthly <- all_vists %>% 
  mutate(
    startOfMonth = floor_date(as_date(mdy_hm(AppointmentDate, tz = NULL)), 'month') 
    # Floors each date to the beginning of the month and adds a new column.
  ) %>% 
  group_by(
    startOfMonth # Groups each entity by the start date of the month
  ) %>% 
  summarise(Appointments = n()) # Aggregates all appointments

all_vists_monthly_ts <- all_vists_monthly %>%
  pull(Appointments) %>% 
  ts(start = c(2017,6), frequency = 12)

all_vists_monthly_ts

which.min(all_vists_monthly_ts)


all_vists_monthly_ts[35] 

round(mean(all_vists_monthly_ts[c(33:34,36)]),0)


```


### Naive Forecast

```{r naiveForecast}

all_vists_monthly_ts_lag_1 <- ts(lag(as.vector(all_vists_monthly_ts)), start = c(2017,6), frequency = 12)


plot(all_vists_monthly_ts, main = 'Naive Forecast\n(Lag = 1)')
lines(all_vists_monthly_ts_lag_1, col = 'firebrick', lty = 'dashed',add = TRUE)
```

### Determining Error


```{r determiningError}

sum((all_vists_monthly_ts - na.remove(all_vists_monthly_ts_lag_1))^2) #MSE

sqrt(sum((all_vists_monthly_ts - na.remove(all_vists_monthly_ts_lag_1))^2))

```

```{r residualPlot}
plot(as.numeric(all_vists_monthly_ts - na.remove(all_vists_monthly_ts_lag_1)))

naive_forecast_tslm <- tslm(all_vists_monthly_ts ~ all_vists_monthly_ts_lag_1)

b <-naive_forecast_tslm$coefficients

forecast::accuracy(naive_forecast_tslm)

sqrt(sum(na.remove(naive_forecast_tslm$residuals^2))/50)

plot(all_vists_monthly_ts, main = 'Naive Forecast\n(Lag = 1)')
lines(fitted(naive_forecast_tslm), col = 'firebrick', lty = 'dashed')
# lines(b[1] + b[2], col = 'skyblue', lty = 'dashed')

arima(naive_forecast_tslm)
```


```{r SecondModel}
model_two <- tslm(all_vists_monthly_ts ~ trend + I(trend^2))

plot(all_vists_monthly_ts, main = 'Zimbi Trend Model')
lines(model_two$fitted, col = 'firebrick', lty = 'dashed')
accuracy(model_two)

summary(model_two)
```


```{r}
nValid <- 48
nTrain <- length(all_vists_monthly_ts) - nValid
train.ts <- window(all_vists_monthly_ts, start = c(2017,6), end = c(2017,nValid))
valid.ts <- window(all_vists_monthly_ts, start = c(2017, nTrain + 1), end = c(2017, nTrain+nValid))
appointments.lm <- tslm(train.ts ~ trend + I(trend^2))
appointments.lm.pred <- forecast(appointments.lm, h = nValid, level = 0)

plot(appointments.lm.pred, ylim = c(0,200), ylab = 'Appointments', xlab = 'Time', bty = 'l',
xaxt = 'n', xlim = c(2017.5,2021), main = 'Zimbi Dental Appointment Forecast', flty = 2)
# plot(appointments.lm.pred, xlim = )
axis(1, at = seq(2017,2022,1), labels = format(seq(2017,2022,1)))
lines(appointments.lm$fitted, lwd = 3)
lines(valid.ts)

accuracy(appointments.lm.pred)

AIC(appointments.lm)
```

